% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Remove the "review" option to generate the final version.
\usepackage{ACL2023}

% This is not strictly necessary, and may be commented out.
% However, it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

% --- Copied: begin
% Standard package includes
\usepackage{times}
\usepackage{latexsym}

\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{amssymb, amsthm, latexsym}
\usepackage{multirow}

\usepackage{nicefrac}

% \usepackage{minted}
% To submit on arxiv, comment the line above and follow this: https://github.com/gpoore/minted/issues/113#issuecomment-888045507
% \usepackage[finalizecache,cachedir=.]{minted}
\usepackage[frozencache,cachedir=.]{minted}

\newminted{python}{%
    fontsize=\fontsize{8.25pt}{8.25pt}\selectfont,
    % options to customize output of pythoncode
}
% --- Copied: end


% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{\textsc{Petals}: Collaborative Inference and Fine-tuning of Large Models}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a seperate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

\author{
  Alexander Borzunov\thanks{\ \ Equal contribution. Correspondence to:\newline \texttt{borzunov.alexander@gmail.com}}\\
  HSE University, Yandex \\\And
  Dmitry Baranchuk$^*$ \\
  Yandex \\\And
  Tim Dettmers$^*$ \\
  University of Washington\\\AND
  Max Ryabinin$^*$ \\
  HSE University, Yandex \\\And
  Younes Belkada$^*$ \\
  Hugging Face, ENS Paris-Saclay \\\And
  Artem Chumachenko \\
  Yandex \\\AND
  Pavel Samygin \\
  Yandex School of Data Analysis \\\And
  Colin Raffel \\
  Hugging Face
}

\begin{document}
\maketitle
\begin{abstract}
Many NLP tasks benefit from using large language models (LLMs) that often have more than 100 billion parameters. With the release of BLOOM-176B and OPT-175B, everyone can download pretrained models of this scale. Still, using these models requires high-end hardware unavailable to many researchers. In some cases, LLMs can be used more affordably via RAM offloading or hosted APIs. However, these techniques have innate limitations: offloading is too slow for interactive inference, while APIs are not flexible enough for research that requires access to weights, attention or logits. In this work, we propose \textsc{Petals}\footnote{\textsc{Petals} source code and documentation are available at \texttt{\href{https://petals.ml}{https://petals.ml}}} --- a~system for inference and fine-tuning of large models collaboratively by joining the resources of multiple parties. We demonstrate that this strategy outperforms offloading for very large models, running inference of BLOOM-176B on consumer GPUs with $\approx$~1 step per second, which is enough for many interactive LLM applications. Unlike most inference APIs, \textsc{Petals} also natively exposes hidden states of served models, allowing to train and share custom model extensions based on efficient fine-tuning methods.
\end{abstract}

\input{intro}
\input{design}
\input{internals}
\input{discussion}
\input{conclusion}

% \clearpage
% \section*{Limitations}

% The main limitations of our work are related to processing sensitive data in the public swarm, since \textsc{Petals} does not guarantee data privacy and correctness of model outputs in this case. We recommend users working with sensitive data to use only trusted servers or set up an isolated \textsc{Petals} swarm.

% We discuss these limitations in more detail in Appendix~\ref{sect:discussion} and acknowledge that the development of methods for privacy-preserving and secure decentralized inference without performance penalties remains an open research problem.

\section*{Ethics Statement}
This work introduces a general-purpose algorithm for decentralized inference of large models, aiming to simplify access to the latest research in deep learning. Thus, we do not envision any direct negative impacts from our research aside from granting the broader public an ability to interact with LLMs trained on uncurated web-crawled data. However, all models we serve are already in open access and thus can be exposed via APIs or other means.

\section*{Acknowledgements}

The authors thank Zheng-Xin Yong, Ilya Dimov, Yozh, Teven Le Scao, Stas Bekman, and Haokun Liu for helpful discussions. We also thank Teven Le Scao for his help in designing Figure~\ref{fig:overview}. A part of the experiments was conducted on a personal server of Elena Voita.

\clearpage

% Entries for the entire Anthology, followed by custom entries
\bibliography{anthology,custom}
\bibliographystyle{acl_natbib}

\end{document}
